### 2019-08-22 chapters/3-background.tex
Add future work

\chapter{Background}\label{chap:background}
\section{JavaScript} \label{sec:background-javascript}
JavaScript (abbreviated JS) is the standard programming language for developing client-side web applications. Together with HTML and CSS, they conform the key technology stack of the Web.

JavaScript is compliant with the ECMAScript Language Specification \citep{ecma-script}. It is an interpreted, dynamically typed language that allows to program small scripts very fast.  Its low learning curve, the possibility of fast prototyping and its easy interaction with the browser's DOM made it the ideal programming languages for introducing dynamic behavior to web applications.

@@

\begin{code}
	\htmlcode{code/background/javascript/jquery-example.html}
	\captionsetup{aboveskip=0pt, belowskip=10pt}
	\caption[Example using JavaScript and HTML]{\textbf{Example using JavaScript and HTML} - Very simple example that uses library jQuery to hide an element from the DOM when it gets clicked. The example was extracted from W3Schools website.}
	\label{code:background-jquery-example}
\end{code}

@@

\figref{fig:background-survey-programming-languages} shows that it is the most popular programming languages among professional developers. Additionally, 5 of the 12 most popular frameworks used for web developing are JavaScript frameworks. Similarly, it can be seen in \figref{fig:background-programming-languages-evolution} that JavaScript's popularity remained constant in the last 3 years while being the most popular programming language for seven consecutive years.

@@

It must be noted that functions and arrays are not built-in types. Functions are considered to be `member of the Object type that is an instance of the standard built-in Function constructor and that may be invoked as a subroutine' \citep{ecma-script}. The \mintinline{text}{Function} object is a built-in object with a \mintinline{text}{[[Call]]} internal property, which means that it can be invoked. Similarly, arrays are instances of the built-in \mintinline{text}{Array} object. Both concepts are shown in \coderef{code:background-functions-and-arrays}.

@@

\subsection{Wrapper Objects for Primitives}
JavaScript also includes the built-in objects \mintinline{text}{Boolean}, \mintinline{text}{String} and \mintinline{text}{Number}. Whenever a method or a property of a primitive value of type \mintinline{text}{boolean}, \mintinline{text}{string} and \mintinline{text}{number} is accessed, JavaScript will convert the primitive value into its corresponding built-in object before accessing the property or invoking the method, as illustrated in \coderef{code:background-javascript-wrapper-objects}. It is however considered a bad practice to explicitly use the built-in objects instead of the primitive values. Several JavaScript linters like JSHint will consider this an error \citep{jshint-wrapper-objects-error}.

@@

\subsection{TypeCoercion} \label{sec:background-js-type-coercion}
Type coercion is the process of converting one type into another one. In JavaScript, a variable can be converted into different types depending on the operator and the value of the other operands. The specific definition can be found in the ECMAScript Language Specification \citep{ecma-script}. The following section will cover the basic forms of type coercion but providing a detailed explanation of how JavaScript converts a variable type into another one for each operator is beyond the scope of this section. It is however intended to give insight about this kind of behavior to show its importance while performing type inference in JavaScript. As a result, specific operator related type coercion is explained only for operator \mintinline{text}{+}.

Finally, there are excellent references regarding JavaScript's Type Coercion, like Kyle Simpson's book `You Don't Know JS: Types \& Grammar' \citep{you-dont-know-js}.

\subsubsection{Types of conversion}
\label{types_of_conversion}
There are two types of type coercion: explicit and implicit \citep{you-dont-know-js}. Explicit type coercion is simply when the developer chooses to convert one type into another one. The developer achieves this by writing a piece of code that shows her intention, as shown in \coderef{code:background-explicit-type-coercion}. This type of coercion is also called type casting \citep{you-dont-know-js}.

@@

\begin{code}
	\jscode{code/background/javascript/type-coercion/implicit-type-coercion.js}
	\captionsetup{aboveskip=0pt, belowskip=10pt}
	\caption[Implicit JavaScript Type Coercion]{\textbf{Implicit JavaScript Type Coercion} - Examples given by Douglas Crockford in his talk `JavaScript: The Good Parts' at Google\footnote{https://www.youtube.com/watch?v=hQVTIJBZook}.}
	\label{code:background-implicit-type-coercion}
\end{code}

@@

The operation \mintinline{text}{ToPrimitive()} is also called within the operators. However, each operator will call either \mintinline{text}{ToString()}, \mintinline{text}{ToNumber()} or \mintinline{text}{ToBoolean()} after calling \mintinline{text}{ToPrimitive()}. This operator will be explained in an independent section.

@@

\begin{code}
	\jscode{code/background/javascript/type-coercion/string-conversion.js}
	\captionsetup{aboveskip=0pt, belowskip=10pt}
	\caption[ToString operation]{\textbf{ToString operation}}
	\label{code:background-to-string-operation}
\end{code}

\subsubsection{ToNumber}
\coderef{code:background-to-number-operation} presents this type of conversion. Strings and boolean values get converted in the expected way. Moreover, \mintinline{text}{undefined} will be converted into \mintinline{text}{NaN} and \mintinline{text}{null} into \mintinline{text}{0}. Similarly, objects are converted into numbers by calling \mintinline{text}{ToPrimtive()} with \mintinline{text}{hint = number}, as presented in \coderef{code:background-to-number-implementation}.

@@

\begin{code}
	\jscode{code/background/javascript/type-coercion/number-conversion.js}
	\captionsetup{aboveskip=0pt, belowskip=10pt}
	\caption[ToNumber operation]{\textbf{ToNumber operation}}
	\label{code:background-to-number-operation}
\end{code}

@@

\begin{code}
	\jscode{code/background/javascript/type-coercion/boolean-conversion.js}
	\captionsetup{aboveskip=0pt, belowskip=10pt}
	\caption[ToBoolean operation]{\textbf{ToBoolean operation}}
	\label{code:background-to-boolean-operation}
\end{code}

@@

An important aspect that should be considered, is that the type of the return value of \mintinline{text}{ToPrimitive()} does not necessarily have to match the chosen \mintinline{text}{preferredType}. This means concretely that, for example, you can call \mintinline{text}{ToPrimitive(a, number)} and get a \mintinline{text}{string} as a result. Of course, when performing an explicit type conversion, JS will eventually convert whatever \mintinline{text}{ToPrimitive} returns into the correct type. But some operators will perform different implicit type conversions \textit{after} calling \mintinline{text}{ToPrimitive()}. The specific type conversion will then be dependent on the type of the return value of \mintinline{text}{ToPrimitive()} and not on the chosen \mintinline{text}{preferredType}.

@@

\section{TypeScript} \label{sec:background-typescript}
\subsection{Declaration Files} \label{sec:declaration-files-background}
Declaration files contain a typed description of an external JavaScript library's API. It allows the IDE to use code intelligence features like auto-completion and avoids misusing the library by declaring specific interfaces and types required by the library. They are necessary for integrating new TypeScript projects with existing JavaScript libraries that are still developed in plain JavaScript.

>>>>

\chapter{Background}\label{chap:background}
\section{JavaScript} \label{sec:background-javascript}
JavaScript (abbreviated JS) is a programming language that is compliant with the ECMAScript Language Specification \citep{ecma-script}. It is one of the
key technologies for developing client-side web applications, alongside HTML and CSS.

It is an interpreted, dynamically typed language that allows to program small scripts very fast.  Its low learning curve, the possibility of fast prototyping and its simple interaction with the contents of a web page through the DOM (Document Object Model) made it the ideal programming languages for introducing dynamic behavior to web applications.

@@

\begin{code}
	\htmlcode{code/background/javascript/jquery-example.html}
	\captionsetup{aboveskip=0pt, belowskip=10pt}
	\caption[Example using JavaScript and HTML]{\textbf{Example using JavaScript and HTML} - Very simple example that uses library jQuery to hide an element from the DOM when it gets clicked. The example was extracted from W3Schools website \citep{w3schools}.}
	\label{code:background-jquery-example}
\end{code}

@@

\figref{fig:background-survey-programming-languages} shows that it is the most popular programming languages among professional developers. Additionally, 5 of the 12 most popular frameworks used for web developing are JavaScript frameworks. Similarly, it can be seen in \figref{fig:background-programming-languages-evolution} that JavaScript's popularity remained constant in the last 3 years and it remained as the most popular programming language for seven consecutive years.

@@

It must be noted that functions and arrays are not built-in types. A Function is considered to be `member of the Object type that is an instance of the standard built-in Function constructor and that may be invoked as a subroutine' \citep{ecma-script}. The \mintinline{text}{Function} object is a built-in object with a \mintinline{text}{[[Call]]} internal property, which means that it can be invoked. Similarly, arrays are instances of the built-in \mintinline{text}{Array} object. Both concepts are shown in \coderef{code:background-functions-and-arrays}.

@@

\subsection{Wrapper Objects for Primitives}
JavaScript also includes the built-in objects \mintinline{text}{Boolean}, \mintinline{text}{String} and \mintinline{text}{Number}. Whenever a method or a property of a primitive value of type \mintinline{text}{boolean}, \mintinline{text}{string} or \mintinline{text}{number} is accessed, JavaScript will convert the primitive value into its corresponding built-in object before accessing the property or invoking the method, as illustrated in \coderef{code:background-javascript-wrapper-objects}. It is however considered a bad practice to explicitly use the built-in objects instead of the primitive values. Several JavaScript linters like JSHint will consider this an error \citep{jshint-wrapper-objects-error}.

@@

\subsection{TypeCoercion} \label{sec:background-js-type-coercion}
Type coercion is the process of converting one type into another one. In JavaScript, a variable can be converted into different types depending on the operator and the value of the other operands. The specific definition can be found in the ECMAScript Language Specification \citep{ecma-script}. The following section will cover the basics about JavaScript's type coercion. A detailed explanation of how JavaScript converts a variable type into another one for each operator can be seen in the ECMAScript Language Specification \citep{ecma-script}. It is however intended to give insight about this kind of behavior to highlight its importance for performing type inference in JavaScript. Therefore, specific operator related type coercion is explained only for operator \mintinline{text}{+}.

There are excellent references regarding JavaScript's Type Coercion, like Kyle Simpson's book `You Don't Know JS: Types \& Grammar' \citep{you-dont-know-js}.

\subsubsection{Types of conversion}
\label{types_of_conversion}
There are two types of type coercion: explicit and implicit \citep{you-dont-know-js}. Explicit type coercion is simply when the developer chooses to convert one type into another one. The developer achieves this by writing a piece of code that explicitly shows her intention, as shown in \coderef{code:background-explicit-type-coercion}. This type of coercion is also called type casting \citep{you-dont-know-js}.

@@

\begin{code}
	\jscode{code/background/javascript/type-coercion/implicit-type-coercion.js}
	\captionsetup{aboveskip=0pt, belowskip=10pt}
	\caption[Implicit JavaScript Type Coercion]{\textbf{Implicit JavaScript Type Coercion} - Examples given by Douglas Crockford in his talk `JavaScript: The Good Parts' at Google \citep{js-the-good-parts}.}
	\label{code:background-implicit-type-coercion}
\end{code}

@@

The operation \mintinline{text}{ToPrimitive()} is also called within the operators. However, each operator will call either \mintinline{text}{ToString()}, \mintinline{text}{ToNumber()} or \mintinline{text}{ToBoolean()} after calling \mintinline{text}{ToPrimitive()}. This operation will be explained in an independent section.

@@

\begin{code}
	\jscode{code/background/javascript/type-coercion/string-conversion.js}
	\captionsetup{aboveskip=0pt, belowskip=10pt}
	\caption[ToString operation examples]{\textbf{ToString operation examples}}
	\label{code:background-to-string-operation}
\end{code}

\subsubsection{ToNumber}
\coderef{code:background-to-number-operation} presents this type of conversion. Strings and boolean values get converted in the expected way. Curiously, \mintinline{text}{undefined} will be converted into \mintinline{text}{NaN} and \mintinline{text}{null} into \mintinline{text}{0}. Similarly, objects are converted into numbers by calling \mintinline{text}{ToPrimtive()} with \mintinline{text}{hint = number}, as presented in \coderef{code:background-to-number-implementation}.

@@

\begin{code}
	\jscode{code/background/javascript/type-coercion/number-conversion.js}
	\captionsetup{aboveskip=0pt, belowskip=10pt}
	\caption[ToNumber operation examples]{\textbf{ToNumber operation examples} - Every non-printable character is removed from the string before converting it into a number.}
	\label{code:background-to-number-operation}
\end{code}

@@

\begin{code}
	\jscode{code/background/javascript/type-coercion/boolean-conversion.js}
	\captionsetup{aboveskip=0pt, belowskip=10pt}
	\caption[ToBoolean operation examples]{\textbf{ToBoolean operation examples} - Values that are coerced to false are called \textit{falsy} values.}
	\label{code:background-to-boolean-operation}
\end{code}

@@

An important aspect that should be considered, is that the type of the return value of \mintinline{text}{ToPrimitive()} does not necessarily have to match the chosen \mintinline{text}{preferredType}. This means concretely that, for example, \mintinline{text}{ToPrimitive(a, number)} can be called and return a \mintinline{text}{string} as a result. Of course, when performing an explicit type conversion, JS will eventually convert whatever \mintinline{text}{ToPrimitive} returns into the correct type. But some operators will perform different implicit type conversions \textit{after} calling \mintinline{text}{ToPrimitive()}. The specific type conversion will then be dependent on the type of the return value of \mintinline{text}{ToPrimitive()} and not on the chosen \mintinline{text}{preferredType}.

@@

\section{TypeScript} \label{sec:background-typescript}
TypeScript is a programming language developed by Microsoft that compiles to plain JavaScript. Its syntax is a superset of JavaScript, which means in practice that `every JavaScript program is also a TypeScript program' \citep{typescript}. It provides optional types annotations that enable features like auto-completion and static checking. Type annotations are erased and variable names are preserved after compiling, which produces a JavaScript output that is very similar to the TypeScript source code. The extension for TypeScript files is \mintinline{text}{.ts}. An example about some of the basic TypeScript features is provided in \coderef{code:background-typescript-and-javascript}.

A complete specification of the language can be found in the TypeScript Language Specification document \citep{typescript}. This section will only cover the most relevant aspects of the language.

\begin{code}
	\begin{bashinline}
$ cat index.js 
function f(a, b) {
	return a + b;
}

f(1, 2);

$ npm i -g typescript
$ mv index.js index.ts
$ tsc index.ts
$ cat index.js
function f(a, b) {
    return a + b;
}
f(1, 2);

$ cat index-with-type-annotations.ts
function f(a: number, b: number) : number {
	return a + b;
}

f(1, 2);

$ tsc index-with-type-annotations.ts
$ cat index-with-type-annotations.js
function f(a, b) {
    return a + b;
}
f(1, 2);
	  \end{bashinline}
	\caption[TypeScript compilation example]{\textbf{TypeScript compilation example} - A plain JavaScript file is renamed into a TypeScript file. The TypeScript compiler will compile it into plain JavaScript again. Secondly, a \mintinline{text}{.ts} file with typed annotations gets compiled into a common JavaScript file with no annotations.}
	\label{code:background-typescript-and-javascript}
  \end{code}

\subsection{Types}
The type analysis performed by TypeScript will only happen at compile-time, which means that the compiled JavaScript has no additional overhead at run-time.

\subsubsection{Any Type}
The \mintinline{text}{any} type closely resembles a common value in JavaScript. It has been introduced to support values with no type annotations. It can be used to represent any JavaScript value. Minimal static checking is performed on values of the type \mintinline{text}{any}: properties with any name can be accessed, methods with any name and any argument list can be invoked and the value can be invoked as a function.

\subsubsection{Primitive Types}
As stated in the specification, `The primitive types are the \mintinline{text}{Number}, \mintinline{text}{Boolean}, \mintinline{text}{String}, \mintinline{text}{Symbol}, \mintinline{text}{Void}, \mintinline{text}{Null} and \mintinline{text}{Undefined} types and all user defined enum types.' \citep{typescript}.

The types \mintinline{text}{Number}, \mintinline{text}{Boolean}, \mintinline{text}{String}, \mintinline{text}{Symbol}, \mintinline{text}{Null} and \mintinline{text}{Undefined} closely match the JavaScript equivalent types. The \mintinline{text}{Void} type stands for the absence of a value, for a example when a function returns no value.

\subsubsection{Object Types}
The following types are considered to be object types: Class and interface type references, array types, tuple types, function types, and constructor types.

Array types can be written as \mintinline{text}{string[]} or as \mintinline{text}{Array<string>}.

The signature of a function can be represented with the \mintinline{text}{Function} type. It is useful for the very common case where a callback is an argument of another function. For example the following function has two arguments, a \mintinline{text}{string} and a callback that receives a \mintinline{text}{number} and returns a \mintinline{text}{boolean}: \mintinline{text}{function f(a: string, c: (x: number) => boolean) { // ... }}.

Interfaces can be used to parametrize object types by defining the types and signatures of properties and methods of a value. It is important to mention that interfaces are not exported in any way to the compiled JavaScript. They are only useful for static type checking at compile-time and code intelligence features provided by the IDE.

Classes are implemented in the expected way according to the usual Object Oriented programming languages. \figref{fig:background-typescript-class-example} shows an example of a TypeScript class implementation and its corresponding JavaScript file after compilation.

\input{figures/background/typescript/classes/class-example.tex}

\subsubsection{Union \& Intersection Types}
A Union Type represents a value that may have only one of multiple possible types. It is represented as \mintinline{text}{let a: string | number}, meaning that \mintinline{text}{a} is either of type \mintinline{text}{string} or \mintinline{text}{number}.

On the other hand, the Intersection Type stands for a value that can simultaneously two types. It can be useful for defining a value that implements two interfaces simultaneously.

\subsection{Declaration Files} \label{sec:declaration-files-background}
Declaration files contain a typed description of an external JavaScript library's API. It allows the IDE to use code intelligence features like auto-completion and avoids misusing the library by declaring specific interfaces and types required by the library. They are necessary for integrating new TypeScript projects with existing JavaScript libraries that are still developed in plain JavaScript.

### 2019-08-22 chapters/6-conclusions.tex
Add future work

Building and end-to-end solution for the generation of TypeScript declaration files was prioritized over type inference accuracy. Consequently, types were taken over from the values at run-time. Obtaining the types from the code examples extracted from the repositories proved to be a pragmatic and effective approximation, enabling to work on specific aspects regarding the TypeScript declaration file generation.

We built a mechanism to automatically create declaration files for every module uploaded to DefinitelyTyped and evaluate the results, creating a TypeScript declaration files parser and a comparator.

We exposed the fundamental aspect of capturing the developer's intention when inferring types in JavaScript. Instead of applying constraints and restrictions for operations with certain types, we presented a proposal where common practices are favored. Uncommon usage is not forbidden but greatly disfavored. Accordingly, we collected evidence regarding the usage of JavaScript operators by analyzing 400 libraries. Furthermore, used in a first place merely as an initial code base for executing the library and approximating types, code examples turned out to be extremely useful for inferring types. They expose the developer's intention explicitly through code.

@@

\section{Future Work} \label{sec:conclusions-future-work}

\subsection{Instrumentation}
\draft{Babel}

\subsection{Declaration files generation}
\draft{Optional parameters}
\draft{Functions as paramters}
\subsubsection{Uploading Results to Definitely Typed Repository}

\subsection{Type Inference}
% \todo{Deberia ser hecho en otro modulo que no sea el runtime information. El runtime information deberia exponer unicamente los datos que le hacen falta al otro para inferir los tipos.}

% \todo{Tambien se puede identificar patterns de programacion donde siempre los tipos son los mismos!!! El concepto de inferir el tipo en realidad pasa a detectar la intencion del developer.}

\subsubsection{Return values}
% \todo{El return value se podria incluir en el analysis de Jalangi, detectando las interactions que tiene etc etc. Esto no es tan facil porque no se puede simplemente inspeccionar el return value, sino que hay que saber para qué se usa. Los ejemplos podrían ser útiles}.

\subsubsection{Variable Names}
% \todo{Se puede extraer info del nombre de las variables y del nombre de los metodos, etc. La developer's intention.}

\subsubsection{Variable Content}

\subsubsection{Function Names}

\draft{JAVERT!!}

>>>>

Building an end-to-end solution for the generation of TypeScript declaration files was prioritized over type inference accuracy. Consequently, types were taken over from the values at run-time. Since developers expose through code how a library should be used, obtaining the types from the code examples extracted from the repositories proved to be a pragmatic and effective approximation, enabling to work on specific aspects regarding the TypeScript declaration file generation itself.

We built a mechanism to automatically create declaration files for potentially every module uploaded to DefinitelyTyped. We managed to generate declaration files for 244 modules. We compared the results against the corresponding files uploaded to DefinitelyTyped by creating a TypeScript declaration files parser and a comparator.

We exposed the fundamental aspect of capturing the developer's intention when inferring types in JavaScript. Instead of applying constraints and restrictions for operations with certain types, we presented a proposal where common practices are favored. Uncommon usage is not forbidden but greatly disfavored. Accordingly, we collected evidence regarding the usage of JavaScript operators by analyzing 400 libraries.

@@

\section{Future Work} \label{sec:conclusions-future-work}
\subsection{Instrumentation}
The latest commit to Jalangi's Github repository was on April 5, 2017 \citep{jalangi-repository}. The are 49 open issues but the last comment on an issue was on January 31, 2018. Jalangi plays a key role in the process of generating declaration files. In order to increase the number of modules that are able to be instrumented shown in \figref{fig:experiments-overall-funnel}, it is necessary to perform some changes to Jalangi's code. Since it is a library that seems to be no longer maintained, the repository might need to be forked.

On of the main drawbacks of Jalangi is that is does not offer full support of ES6, as stated in their repository: `Jalangi2 supports ECMAScript 5.1. Some ES6 features may work, but have not been tested.' \citep{jalangi-repository}. A possible solution for this would be to use Babel to convert ES6 code into a backwards compatible version like ECMAScript 5.1 before instrumenting the code \citep{babel}.

\subsection{Optional parameters}
The concept of optional parameters was not implemented. The difference between an optional parameter and a parameter of type \mintinline{text}{undefined} has to be further studied. In practice, JavaScript does return \mintinline{text}{undefined} for both the cases of accessing a nonexistent property or a property with the value \mintinline{text}{undefined}. Therefore, the developer's intention of treating a parameter as optional has to be clearly captured.

\subsection{Function types}
TypeScript enables the developer to clearly specify the signature of a function that is a parameter of an interface or of another function. This is considered a good practice and it greatly enriches the declaration file. The current implementation assigns to a parameter that is a function the type \mintinline{text}{Function}, which is naturally very unspecific. The gathered run-time information is already sufficient for creating the signatures of function types. The code that uses the run-time information for clearly representing the signatures needs to be developed.

\subsection{Uploading Results to DefinitelyTyped}
Automatic creation of pull requests on the DefinitelyTyped repository could be attached to the pipeline. Compliance with the contribution guidelines specified by DefinitelyTyped can be checked before generating the pull requests.

\subsection{Type Inference}
Inspecting the semantics of the code can be of great use for inferring types. The name that a developer chooses for a variable or a function is very important to determine the purpose of that variable. A function that performs the operation \mintinline{text}{firstArgument + secondArgument} has a complete different meaning if its called \mintinline{text}{sum()} or \mintinline{text}{concatenate()}, even though the JavaScript operator is the same in both cases.

Furthermore, the content of a value can also be useful. Some strings may represent, for example, known error codes or RGB colors representations. This information can be used by a type inferrer to determine the type more accurately.

Finally, detecting common development patterns is important. Developers tend to apply the same patterns for solving similar problems. For example, a very common pattern that developers use for dealing with optional parameters is \mintinline{text}{let a = firstArgument || {}}. Identifying such patterns and relating them with commonly used types can be extremely useful for a better capture of the developer's intention.
